The state of knowledge for marine invertebrate biodiversity in the continental US

### 2018-01-16 manuscript.Rmd
edit ITIS numbers and discussion

This underrepresentation is also manifest in the intensity of the collecting effort. In OBIS records for the US EEZ, marine invertebrates represent `r invert_props$prop_obis_invert_samples`% of the occurrences but `r invert_props$prop_obis_invert_species`% of the total number of species; and in iDigBio, they represent `r invert_props$prop_idigbio_invert_samples`% of the records but `r invert_props$prop_idigbio_invert_species`% of the species. In contrast, marine vertebrates represent `r invert_props$prop_idigbio_vert_samples`% of the samples and `r invert_props$prop_idigbio_vert_species`% of the species in iDigBio, and `r invert_props$prop_obis_vert_samples`% of the samples and `r invert_props$prop_obis_vert_species`% of the species in OBIS (Fig. \@ref(fig:kingdom-diversity-plot)). In other words, while marine invertebrates are represented at a level on par with their diversity in the marine fraction of the databases, each species is represented by fewer samples compared to vertebrates (median number of records per species for invertebrates `r median_idigbio_invert` and `r median_obis_invert`, _vs_ and `r median_idigbio_vert` and `r median_obis_vert` for vertebrates, for iDigBio and OBIS respectively).

```{r itis-stats}
prop_molluscs <- itis_molluscs_in_dbs %>%
    dplyr::summarize(
               prop_idigbio_us = format_output(mean(n_idigbio_geo_us > 0,
                                                    na.rm = TRUE)* 100),
               prop_obis_us = format_output(mean(n_obis_geo_us > 0,
                                                 na.rm = TRUE) * 100)
           )

prop_crustaceans <- itis_crustaceans_in_dbs %>%
    dplyr::summarize(
             prop_idigbio_us = format_output(mean(n_idigbio_geo_us > 0,
                                                  na.rm = TRUE) * 100),
             prop_obis_us = format_output(mean(n_obis_geo_us > 0,
                                             na.rm = TRUE) * 100)               
           )
```

To assess the fraction of digitization of the fauna, we compared the species included in iDigBio and OBIS with the comprehensive checklists for molluscs [@turgeon1998] and crustaceans [@mclaughlin2005] that are available for the United States. For molluscs, of the `r format(nrow(itis_molluscs_in_dbs), big.mark=",")` species included in the checklist, `r prop_molluscs$prop_idigbio_us`% are in iDigBio and `r prop_molluscs$prop_obis_us`% are in OBIS. For crustaceans, of the `r format(nrow(itis_crustaceans_in_dbs), big.mark=",")` species included in the checklist, `r prop_crustaceans$prop_idigbio_us`% are in iDigBio and `r prop_crustaceans$prop_obis_us`% are in OBIS.

@@

>>>>

@@

This underrepresentation is also manifest in the intensity of the collecting effort. Marine invertebrates are estimated to represent  `r total_invert_prop`% of the total marine diversity[@worms-stats]. In OBIS records for the US EEZ, marine invertebrates represent `r invert_props$prop_obis_invert_samples`% of the occurrences but `r invert_props$prop_obis_invert_species`% of the total number of species; and in iDigBio, they represent `r invert_props$prop_idigbio_invert_samples`% of the records but `r invert_props$prop_idigbio_invert_species`% of the species. In contrast, marine vertebrates represent `r invert_props$prop_idigbio_vert_samples`% of the samples and `r invert_props$prop_idigbio_vert_species`% of the species in iDigBio, and `r invert_props$prop_obis_vert_samples`% of the samples and `r invert_props$prop_obis_vert_species`% of the species in OBIS (Fig. \@ref(fig:kingdom-diversity-plot)). In other words, while marine invertebrates are represented at a level on par with their diversity within the US EEZ in the databases, each species is represented by fewer samples than for vertebrates (median number of records per species for invertebrates `r median_idigbio_invert` and `r median_obis_invert`, _vs._ and `r median_idigbio_vert` and `r median_obis_vert` for vertebrates, for iDigBio and OBIS respectively).

To assess the level of digitization of the fauna, we compared the species included in iDigBio and OBIS with the comprehensive checklists for molluscs [@turgeon1998] and crustaceans [@mclaughlin2005] that are available for the United States. For molluscs, of the `r format(itis_stats$moll_nspp, big.mark=",")` species included in the checklist, `r itis_stats$prop_moll_idig`% are in iDigBio and `r itis_stats$prop_moll_obis`% are in OBIS. For crustaceans, of the `r format(itis_stats$crust_nspp, big.mark=",")` species included in the checklist, `r itis_stats$prop_crust_idig`% are in iDigBio and `r itis_stats$prop_crust_obis`% are in OBIS. Given that molluscs and crustaceans are among the most sampled taxa, and that they have received more taxonomic attention than other phyla (see below), these percentages offer a conservative estimate for the fraction of the fauna that is available digitally. Therefore, we can estimate that there is at least 28,000 species of marine invertebrates in US waters. The higher overlap between the comprehensive checklist for  crustaceans and OBIS is attributable to occurences of planktonic species that are more represented in OBIS data than in museum collections.
